---
title: "Batch processing with eyeQuality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Batch processing with eyeQuality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eyeQuality)
```

`vignette("getting-started")` walks through preprocessing a single recording
with `eyeQuality()`. Most real projects have more than one recording, though
-- one file per participant, or per participant per session -- and running
`eyeQuality()` by hand for each one gets tedious fast. `eyeQualityBatch()`
finds every raw recording under a directory tree, runs `eyeQuality()` on each
of them (in parallel, across multiple cores), and writes the same set of
output files `eyeQuality(saveData = TRUE)` would write for a single file,
once per recording, plus one batch-level summary describing the whole run.

## Building a small example dataset

`eyeQualityBatch()` needs an actual directory of files to point at, so this
vignette builds a small one from the same synthetic fixture
`vignette("getting-started")` uses for a single file -- copied four times into
a two-subject, two-session layout, entirely inside a temporary directory. This
is only to give the examples below something real to run against; nothing
here is specific to Tobii Studio exports, and the file names below follow a
plain `sub-<id>_ses-<id>_..._physio.tsv` convention, not a requirement
`eyeQualityBatch()` itself imposes.

```{r}
raw_file <- system.file("extdata", "tobii_studio_sample.tsv", package = "eyeQuality")

batch_dir <- file.path(tempdir(), "eyeQuality-batch-demo")
dir.create(batch_dir, showWarnings = FALSE)

for (subject in c("sub-01", "sub-02")) {
  for (session in c("ses-01", "ses-02")) {
    session_dir <- file.path(batch_dir, subject, session)
    dir.create(session_dir, recursive = TRUE, showWarnings = FALSE)
    file.copy(
      raw_file,
      file.path(session_dir, paste0(subject, "_", session, "_task-test_recording-eyetracking_physio.tsv"))
    )
  }
}

list.files(batch_dir, recursive = TRUE)
```

## How eyeQualityBatch() finds files: listBidsFiles()

Before dispatching any work, `eyeQualityBatch()` calls `listBidsFiles()` to
build the list of raw files to process. `listBidsFiles()` supports two
`layout` modes, matched to how eye-tracking datasets tend to actually be
organized:

- **`layout = "bids"`** (the default) -- a fixed two-level
  `sub-<id>/ses-<id>/...` hierarchy, matched by regular expressions
  (`subjectPattern_regex`/`sessionPattern_regex`, defaulting to
  `"sub-[A-Z0-9]+"`/`"ses-[0-9]+"`). This is the layout the example directory
  above already follows:

  ```{r}
  listBidsFiles(batch_dir)
  ```

  A directory that doesn't match the subject/session pattern is skipped
  rather than silently descended into or erroring out the whole batch; every
  skipped path is reported in `attr(result, "skipped")` so you can confirm
  nothing was dropped by mistake.

- **`layout = "glob"`** -- for directory structures that don't fit that fixed
  two-level shape. Instead of subject/session regexes, you supply a
  glob-style `pathPattern` (matched relative to `directory`, using `/`
  regardless of OS): `*` matches any characters within one path segment,
  `?` matches a single character within a segment, and `**` matches zero or
  more whole segments -- so `"**/*.tsv"` finds `.tsv` files at any depth,
  while `"*/*.tsv"` matches only files exactly one directory level down. An
  optional `excludePattern_regex` drops matched paths back out of the result
  (into `attr(result, "skipped")`, same as the `"bids"` skip reporting) --
  useful when a participant's own processed-output subfolder would otherwise
  be re-ingested as raw data:

  ```{r}
  listBidsFiles(
    batch_dir,
    layout = "glob",
    pathPattern = "**/*.tsv",
    excludePattern_regex = "derivatives"
  )
  ```

  `layout = "glob"` covers two directory shapes that don't fit the BIDS-like
  default: a **paired-in-place** layout, where each participant's folder
  holds their raw file(s) and processed output together rather than in a
  separate `derivatives/` subtree (set `pathPattern` to match the raw files
  at whatever depth they sit, and `excludePattern_regex` to skip each
  participant's own output subfolder), and a **centralized-output** layout,
  where every participant's processed output lands in one shared directory
  outside the per-participant raw folders (point `directory` at the raw-data
  root and, if the shared output directory happens to live inside that same
  root, exclude it by name).

`eyeQualityBatch()` accepts the same arguments `listBidsFiles()` does
(`layout`, `subjectPattern_regex`/`sessionPattern_regex`, `pathPattern`,
`excludePattern_regex`, `modalityPattern_regex`, `recursiveSearch`) via its
`...`, so any of the calls above work as-is inside an `eyeQualityBatch()`
call too.

## Running a batch

`eyeQualityBatch()` needs a `directoryBIDS` (despite the name, this is just
"the directory to search" -- it works with either `layout`) and a
`batchName`, which gets stamped into every output filename so you can tell
one run apart from another. Display dimensions and any other `eyeQuality()`
argument are passed straight through and applied to every file in the batch:

```{r, results = "hide", message = FALSE}
eyeQualityBatch(
  batch_dir,
  batchName = "demo",
  numberCores = 1,
  displayDimensionX_mm = 594,
  displayDimensionY_mm = 344
)
```

(`numberCores` defaults to roughly 85% of the machine's available cores,
capped at one core per file; it's pinned to `1` above purely so this
vignette's timing is predictable, not because that's a recommended setting.)

Unlike `eyeQuality()`'s single-file progress log, `eyeQualityBatch()` doesn't
print its run log to the console -- with multiple files potentially
processing across several worker processes at once, an interleaved console
stream wouldn't be very readable anyway. Instead, everything is written to a
`preprocessing_batch_summary_desc-<batchName>.txt` file at the root of the
directory you pointed it at:

```{r}
summary_file <- file.path(batch_dir, "preprocessing_batch_summary_desc-demo.txt")
cat(readLines(summary_file), sep = "\n")
```

Per-recording output -- the same `events`/`preproc`/`preproc_runtimes`/
`preproc_qcsummary` files and run log `vignette("getting-started")` describes
for a single `eyeQuality(saveData = TRUE)` call -- lands next to each raw
file, under a `derivatives/eyeQuality-v1/` subdirectory by default (or
wherever `outputDir` points, if you set it):

```{r}
list.files(batch_dir, recursive = TRUE, pattern = "desc-demo")
```

### Reading the batch summary back in

`parsePreprocessingBatchSummary()` parses that summary file back into R
rather than requiring you to read it as text. `info_to_extract = "summary"`
gets the one-row run summary (file counts, data size, timing):

```{r}
parsePreprocessingBatchSummary(summary_file, info_to_extract = "summary")
```

`"successfulfiles"` and `"skippedfiles"` return a plain character vector of
filepaths; `"failedfiles"` returns a `file`/`error` data frame, covered
below.

## Resuming an interrupted or partially completed batch

Batch runs over real datasets can be large enough that re-running everything
from scratch after an interruption -- or after adding a few new
participants -- is wasteful. Before dispatching any file,
`eyeQualityBatch()` checks whether a matching `preproc_qcsummary` output
already exists for that file under the given `batchName` (and `outputDir`,
if set); if it does, the file is skipped rather than reprocessed. Calling
`eyeQualityBatch()` again against the same directory and `batchName`, with
nothing changed, finds every file already done and skips all of them:

```{r, results = "hide", message = FALSE}
eyeQualityBatch(
  batch_dir,
  batchName = "demo",
  numberCores = 1,
  displayDimensionX_mm = 594,
  displayDimensionY_mm = 344
)
```

```{r}
parsePreprocessingBatchSummary(summary_file, info_to_extract = "skippedfiles")
```

This is also what makes it safe to add new files to an in-progress dataset
and re-run the same call: only the new files (the ones without an existing
`preproc_qcsummary` output yet) get dispatched, and already-completed
files' output is left untouched rather than overwritten. If you do want to
force every candidate file to be reprocessed regardless of existing output
-- for example, after changing a preprocessing argument and wanting it
applied everywhere -- pass `force = TRUE`:

```{r, results = "hide", message = FALSE}
eyeQualityBatch(
  batch_dir,
  batchName = "demo",
  numberCores = 1,
  displayDimensionX_mm = 594,
  displayDimensionY_mm = 344,
  force = TRUE
)
```

## Structured failure reporting

Not every file in a large batch will make it through preprocessing --
a malformed export, an unsupported column layout, or a genuine data problem
can all cause `eyeQuality()` to raise an error for a given file. Rather than
stopping the whole batch, `eyeQualityBatch()` catches that error, lets the
rest of the batch continue, and records the failure -- both the file and its
error message -- in the summary file's "Files that failed processing"
section.

To see this in action, this adds one more file to the example directory
whose timestamps are deliberately out of order (a real, if artificial, cause
of a preprocessing failure) and re-runs the batch:

```{r, results = "hide", message = FALSE}
broken <- read.delim(raw_file, check.names = FALSE)
# swap two rows' timestamps so they're no longer chronologically ordered
broken[c(5, 6), c("EyeTrackerTimestamp", "RecordingTimestamp")] <-
  broken[c(6, 5), c("EyeTrackerTimestamp", "RecordingTimestamp")]

broken_session_dir <- file.path(batch_dir, "sub-03", "ses-01")
dir.create(broken_session_dir, recursive = TRUE, showWarnings = FALSE)
write.table(
  broken,
  file.path(broken_session_dir, "sub-03_ses-01_task-test_recording-eyetracking_physio.tsv"),
  sep = "\t",
  row.names = FALSE,
  quote = FALSE
)

eyeQualityBatch(
  batch_dir,
  batchName = "demo",
  numberCores = 1,
  displayDimensionX_mm = 594,
  displayDimensionY_mm = 344
)
```

```{r}
parsePreprocessingBatchSummary(summary_file, info_to_extract = "failedfiles")
```

The rest of the batch -- every file that already succeeded -- is unaffected;
only the newly added file shows up as failed, with the same diagnostic
message `eyeQuality()` would have raised if you'd run it on that one file
directly. A subsequent `eyeQualityBatch()` call against the same directory
and `batchName` (still without `force = TRUE`) will retry a previously
failed file rather than treating it as done, since a failed run never
produces the `preproc_qcsummary` output the resumability check looks for.

```{r, include = FALSE}
unlink(batch_dir, recursive = TRUE)
```

## Where to go next

- `?eyeQualityBatch`, `?listBidsFiles`, and `?parsePreprocessingBatchSummary`
  document every argument in full.
- `?eyeQuality-schema` documents the columns each per-recording output file
  contains.
- `vignette("getting-started")` covers preprocessing a single recording, and
  the arguments (`eyeSelection_method`, `validityThreshold`,
  `smoothGaze_boolean`, and so on) that `eyeQualityBatch()` passes through to
  every file in a batch.
