---
title: "Getting started with eyeQuality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with eyeQuality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(eyeQuality)
```

## What eyeQuality does

**eyeQuality** takes raw exports from eye tracking software and turns them
into a standardized, quality-checked dataset: gaze positions in both pixel
and visual-angle space, pupil size, blink and fixation/saccade
classification, and a set of data-quality metrics describing how much of the
recording was usable. The goal is to make preprocessing transparent and
reproducible -- every step the pipeline takes is a named, individually
testable function, and the same call works whether your data came from Tobii
Studio or Tobii Pro.

This vignette walks through preprocessing a single recording end to end. It
uses one of the small synthetic fixture files bundled with the package
(`system.file("extdata", ..., package = "eyeQuality")`), so every chunk below
runs as-is -- copy it into your own session to follow along.

## Supported eye trackers

eyeQuality doesn't ask you which device produced a file. Instead, it keeps an
internal *registry* of eye tracker adapters -- one adapter per device/export
format -- and asks each registered adapter in turn "does this look like your
data?" The first adapter that recognizes the file's raw column layout handles
the rest of the import.

```{r}
names(registered_adapters())
```

Out of the box, eyeQuality recognizes exports from **Tobii Studio** and
**Tobii Pro**. Each adapter also declares a `geometry_type` -- both of the
built-in adapters are `"screen"`-based recordings (a participant looking at a
fixed monitor, as opposed to a head-mounted tracker):

```{r}
vapply(registered_adapters(), function(a) a$geometry_type, character(1))
```

The registry is also how eyeQuality is meant to be extended: a new device
gets its own adapter -- five small functions that translate its raw export
into eyeQuality's common column schema (`?eyeQuality-schema`) -- registered
with `register_eyetracker_adapter()`. See `?new_eyetracker_adapter` for the
interface each adapter must implement if you're adding support for a device
that isn't built in.

## Preprocessing a single recording

The main entry point is `eyeQuality()`. At minimum it needs a file path and
the physical dimensions of the display the data was recorded on, **in
millimeters** -- these are what let the pipeline convert pixel-space gaze
positions into visual angle later on.

```{r}
raw_file <- system.file("extdata", "tobii_studio_sample.tsv", package = "eyeQuality")
raw_file
```

```{r, results = "hide", message = FALSE}
result <- eyeQuality(
  raw_file,
  displayDimensionX_mm = 594, # e.g. a 1920x1080 display, physically 594mm wide
  displayDimensionY_mm = 344, # ...and 344mm tall
  saveData = FALSE
)
```

`eyeQuality()` prints a running log of every pipeline stage as it works
(import, adapter detection, validity/offscreen masking, interpolation, blink
and fixation/saccade classification, and so on) -- useful when you're
diagnosing a single problem file, but noisy for a vignette, so it's
suppressed above. Set `saveData = TRUE` and the same log is written to a run
log file alongside your output instead of printing to the console (see
"Saving results to disk" below).

With `saveData = FALSE` (the default), `eyeQuality()` returns the
preprocessed data as a data frame:

```{r}
nrow(result)
colnames(result)
```

This particular fixture holds gaze constant at a single off-center screen
location for its whole 200-sample recording, purely so its expected values
are known exactly -- a real recording will show natural variation, fixations,
saccades, and some missing/blink data. The columns you'll care about most for
downstream analysis are the `.preprocessed`-suffixed ones -- eyeQuality's
final answer for gaze position, pupil size, distance from the screen, and
velocity after every cleaning step has been applied:

```{r}
result[5, c(
  "gazeX.preprocessed_px", "gazeY.preprocessed_px",
  "gazeX.preprocessed_va", "gazeY.preprocessed_va",
  "distanceZ.preprocessed_mm", "pupil.preprocessed",
  "IVT.classification"
)]
```

`IVT.classification` (and its neighbors `IVT.fixationIndex`,
`IVT.saccadeIndex`, `IVT.fixationDuration_ms`) records what the
velocity-threshold (I-VT) algorithm decided each sample was: a fixation, a
saccade, a blink, or missing/unclassified.

### Key optional arguments

`eyeQuality()` has a number of arguments beyond the required file path and
display dimensions; the ones you're most likely to reach for are:

- `eyeSelection_method` -- how to combine left/right eye data into a single
  gaze stream. `"Maximize"` (the default) uses both eyes when available and
  falls back to whichever eye has data when only one does; `"Strict"` only
  keeps a sample when *both* eyes have valid data; `"Left"`/`"Right"` use
  only that eye.
- `validityThreshold` -- the minimum per-sample confidence (on a generic
  `0`-`1` scale, `1` = highest confidence) to keep. Left `NULL` by default,
  which uses each device's own sensible default rather than a single
  hardcoded cutoff -- device-native validity scales differ (Tobii Studio's
  native `0`-`4` code vs. Tobii Pro's binary valid/invalid flag), and
  eyeQuality converts your `0`-`1` value onto whichever scale the detected
  device actually uses.
- `smoothGaze_boolean` -- whether to apply smoothing before computing visual
  angle and velocity. `TRUE` by default.
- `timeStart`/`timeEnd`, or `studioEvents`/`proEvents` -- restrict
  preprocessing to a specific time window, either by explicit
  `recordingTimestamp_ms` bounds or by a pair of event labels marking the
  start/end of a task.
- `includeIntermediates` -- keep every intermediate column the pipeline
  produces along the way (raw-but-masked `.valid` columns, per-stage
  `.int`/`.eyeSelect`/`.smooth` columns), instead of only the final
  `.preprocessed` ones. Needed if you want to call `calculateOutputMetrics()`
  yourself on the result (see below).
- `verbose` -- when `TRUE`, adds extra diagnostic messages for things worth a
  second look that a quick summary wouldn't surface on its own -- e.g. a
  column that's present but entirely `NA`, or a long run of consecutive
  samples below the validity threshold.

## Quality metrics

Every `eyeQuality()` run also computes a summary table of data-quality
metrics -- percent of raw data missing per eye, percent classified as
blinks/fixations/saccades, robustness ratios, and more -- and prints it to
the console as part of the run. If you want that table back as a data frame
to work with programmatically, call `calculateOutputMetrics()` yourself. It
expects the intermediate columns `eyeQuality()` strips out by default, so ask
for them with `includeIntermediates = TRUE`:

```{r, results = "hide", message = FALSE}
result_full <- eyeQuality(
  raw_file,
  displayDimensionX_mm = 594,
  displayDimensionY_mm = 344,
  saveData = FALSE,
  includeIntermediates = TRUE
)
metrics <- calculateOutputMetrics(result_full)
```

`metrics` is a data frame with one row per quality metric and columns for
count (`n`), proportion (`percent`), and (where the metric is a duration or
similar continuous quantity) summary statistics like `group_mean`/`group_sd`:

```{r}
metrics[c("valid_raw_data", "robustness_proportion_valid_data_to_all_data", "ivt_fixations"), c("n", "percent")]
```

Here, for example, 200/200 raw samples were valid, and 198 of the 200 final
samples (99%) were classified as fixations -- consistent with a recording
that never moves.

## Visualizing a recording

`generateEyeTrackingPlots()` takes a preprocessed data frame and returns a
list of three `ggplot2` plots: (1) raw per-channel diagnostic plots for gaze
X/Y, distance, and pupil size for each eye, (2) a heatmap of gaze density
across the screen, and (3) the final smoothed gaze trace with fixation
regions highlighted.

```{r, warning = FALSE, fig.alt = "Diagnostic plots of raw gaze, pupil, and distance data for each eye"}
plots <- generateEyeTrackingPlots(result)
plots[[1]]
```

(Warnings are suppressed above only because this fixture's gaze never moves,
which makes some `ggplot2` statistics degenerate on a single point; a real
recording won't trigger them.)

## Saving results to disk

Passing `saveData = TRUE` instead of returning the data frame writes four
files: the preprocessed gaze stream, the extracted event stream, the
per-stage run timings, and the quality metrics table -- plus a run log
capturing everything `eyeQuality()` would otherwise print to the console.
By default these are written to a `derivatives/eyeQuality-v1/` subdirectory
next to the input file; pass `outputDir` to write somewhere else instead, and
`batchName` to tag the output files (useful when re-running the same file
with different settings):

```{r}
out_dir <- file.path(tempdir(), "eyeQuality-demo")
```

```{r, results = "hide", message = FALSE}
eyeQuality(
  raw_file,
  displayDimensionX_mm = 594,
  displayDimensionY_mm = 344,
  saveData = TRUE,
  outputDir = out_dir,
  batchName = "demo"
)
```

```{r}
list.files(out_dir)
```

```{r, include = FALSE}
unlink(out_dir, recursive = TRUE)
```

## Processing more than one file

For a whole dataset rather than a single file, `eyeQualityBatch()` runs
`eyeQuality()` across every recording found by `listBidsFiles()` (which
understands both a fixed BIDS-like `sub-*/ses-*/` directory layout and more
flexible glob-style patterns) and writes one set of output files per
recording. See `?eyeQualityBatch` and `?listBidsFiles` for the full set of
options, including how to point it at a YAML batch configuration file
instead of repeating arguments per call.

## Where to go next

- `?eyeQuality-schema` documents every column the pipeline produces or
  consumes, with type, units, and NA semantics.
- `help(package = "eyeQuality")` lists every exported function -- most of the
  individual pipeline stages `eyeQuality()` calls internally (`eyeSelection()`,
  `classifyBlinks()`, `classifyGazeIVT()`, `interpolateGaze()`, and so on) are
  exported and usable on their own if you want to build a custom pipeline
  instead of using the `eyeQuality()` wrapper.
- `?new_eyetracker_adapter` and `?register_eyetracker_adapter` cover adding
  support for a device that isn't built in.
